% ============================================================
% appendix.tex -- Supplementary material
% ============================================================

\appendix

\section{Full MCP tool schema}
\label{app:tools}

Table~\ref{tab:tools} gives the complete list of the 24~MCP tools. The
canonical JSON Schema (with per-argument constraints, return types, and
error codes) lives in \texttt{configs/mcp-tools.json} and is loaded
directly by the FastMCP server.

\begin{table}[h]
\centering
\caption{MCP tool surface. Tool names are camelCase at the MCP
boundary and snake\_case internally.}
\label{tab:tools}
\small
\begin{tabular}{lll}
\toprule
Wave & Type & Tools \\
\midrule
1 & read
    & \texttt{searchEncounters}, \texttt{searchPatients},
      \texttt{searchClinicalKnowledge}, \\
 & & \texttt{searchReferenceMaterials}, \texttt{searchAvailableResources}, \\
 & & \texttt{getEncounterDetails}, \texttt{getConditionDetails},
      \texttt{getPatientHistory}, \\
 & & \texttt{getProtocolDetails}, \texttt{getTransferStatus},
      \texttt{getInsuranceCoverage}, \\
 & & \texttt{getReferenceArticle} \\
\addlinespace
2 & compute
    & \texttt{checkResourceAvailability},
      \texttt{calculateTransferTime}, \\
 & & \texttt{runDecisionRule}, \texttt{validateTreatmentPlan} \\
\addlinespace
3 & mutate
    & \texttt{createClinicalOrder}, \texttt{updateTaskStatus},
      \texttt{updateEncounter}, \\
 & & \texttt{updatePatientRecord}, \texttt{registerPatient},
      \texttt{applyProtocol} \\
\addlinespace
4 & workflow & \texttt{processDischarge}, \texttt{processTransfer} \\
\bottomrule
\end{tabular}
\end{table}

All wave-1 search tools cap results at 10 with no \texttt{hasMore}
signal. All wave-3 and wave-4 tools validate inputs before applying
the mutation and log structured audit entries. Error responses are
uniform:
$\{\texttt{status}{:}\,\texttt{"error"},\ \texttt{code}{:}\,\ldots,\ \texttt{message}{:}\,\ldots\}$.

\section{Criteria by dimension}
\label{app:criteria}

The 2{,}255 binary criteria decompose approximately as follows
across the six weighted dimensions, aggregated over all task
categories: \textsf{clinical\_completeness}~$\sim$20\%;
\textsf{clinical\_correctness}~$\sim$25\%;
\textsf{protocol\_adherence}~$\sim$15\%;
\textsf{documentation\_quality}~$\sim$10\%;
\textsf{safety}~$\sim$20\% (the 515 safety-critical criteria are a
superset of the \textsf{safety} dimension and can additionally appear
in any of the other dimensions when they were authored with
safety\_critical=true);
\textsf{temporal\_sequencing}~$\sim$10\%. Exact counts per category
are emitted by \texttt{scripts/analyze\_results.py}. Dimension weights
are used only for post-hoc error decomposition; they do not enter
Eq.~\ref{eq:reward}.

\section{Pilot v2--v8 changelog}
\label{app:pilots}

The six infrastructure bugs described in \S\ref{sec:v7-superseded} were
fixed across pilots v6, v7, and v8. Full per-task pass/fail tables for
every pilot live in \texttt{results/pilot-v\{3..8\}-\{claude-opus,gpt54\}/}.
Pilot-level deltas for Pass@1 and mean reward across the v3--v8 run
are plotted in Figure~\ref{fig:progression}.

\section{Wilson score confidence interval derivation}
\label{app:wilson}

For $k$ successes out of $n$ trials, the Wilson score 95\% CI is
\[
\left(\frac{\hat p + \frac{z^2}{2n}}{1 + z^2/n}\right)
\pm
\frac{z\sqrt{\frac{\hat p(1-\hat p)}{n} + \frac{z^2}{4n^2}}}{1 + z^2/n},
\]
with $\hat p = k/n$ and $z = 1.96$. Wilson is preferred over the
normal approximation when $n \hat p (1 - \hat p)$ is small; over
Clopper--Pearson it is less conservative without becoming
anti-conservative in the $n\sim 200$ regime relevant for this
benchmark. The implementation is
\texttt{src/metrics/confidence\_intervals.py}.

\section{Judge reliability details}
\label{app:judge}

Cross-vendor assignment is pinned by
\texttt{src/healthcraft/llm/judge.py}:
Claude Opus~4.6 $\to$ GPT-5.4 judge; GPT-5.4 $\to$ Claude Opus~4.6
judge; Gemini~3.1~Pro $\to$ Claude Opus~4.6 judge. Cohen's
$\kappa=0.553$ was computed over a 100-criterion test-retest sample
(each criterion graded three times, $n=300$ calls), with 77.0\% raw
self-agreement. Oracle validation on task CC-001 returned 5/5 agreement
with the \texttt{world\_state} verification method, indicating that for
the \texttt{world\_state}-verifiable subset the judge does not
introduce additional drift.

\paragraph{Stratification by trajectory length.}
Agreement and $\kappa$ were stratified by trajectory length: short
trajectories ($<15$ turns, $n=5$) showed 60.0\% agreement /
$\kappa=0.444$; medium ($15$--$40$ turns, $n=70$) 72.9\% / $0.542$;
long ($>40$ turns, $n=25$) 92.0\% / $0.306$. The long-trajectory
stratum is the paradoxical case: raw agreement is highest there but
$\kappa$ is lowest. This is Cohen's kappa's well-known base-rate
artefact: when agreement is near the ceiling (the judge almost
always returns ``satisfied'' on long trajectories because those
trajectories tend to contain the evidence the criterion is looking
for), the chance-agreement denominator grows faster than the observed
agreement numerator and $\kappa$ collapses. We report both statistics
because they answer different questions: raw agreement is the rate
at which re-grading would flip a verdict, while $\kappa$ is the rate
above chance. The long-trajectory regime is more relevant for Claude
(median $\sim 49$ turns) than GPT-5.4 (median $\sim 8$ turns), which
is why the judge-reliability concern in \S\ref{sec:limits} is framed
as a bound on Claude-favouring variance rather than symmetric noise.

\section{V9 deterministic overlay audit of the LLM judge}
\label{app:v9-overlay}

Cross-vendor judging (\S\ref{sec:eval}) does not eliminate judge
variance; Cohen's $\kappa=0.553$ leaves appreciable residual
disagreement. Because the judge is the unit of measurement for
63.5\% of criteria % CN:v2_llm_judge_pct
(\S\ref{sec:limits}), its noise floor bounds every downstream claim.
We therefore built an independent deterministic audit:
44 criteria % CN:v9_overlay_criteria
originally graded by \texttt{llm\_judge} were migrated to
\texttt{world\_state} predicates and re-graded against the cached V8
trajectories. This appendix reports what that audit found; the full
per-observation report is committed at
\texttt{docs/V9\_OVERLAY\_AUDIT.json}.

\paragraph{Overlay construction.}
The migration pipeline has three stages. (i)~Auto-migration:
\texttt{scripts/migrate\_criteria.py} scans every
\texttt{llm\_judge} assertion for patterns that cleanly map onto
audit-log predicates (e.g.\ ``agent ordered $X$'' $\mapsto$
``audit\_log contains \texttt{createClinicalOrder} with
medication$=X$''). (ii)~Confidence filtering:
\texttt{scripts/populate\_v9\_overlay.py} retains only
\texttt{confidence == 'high'} candidates and flags
subordinate-clause assertions for review.
(iii)~Manual correction of a small number of wrong-entity
mappings (e.g.\ an overlay row whose assertion concerned blood
cultures was originally wired to a blood-product tool). The
resulting overlay
(\texttt{configs/rubrics/v9\_deterministic\_overlay.yaml}) has
$44$ entries across five categories.

\paragraph{The \texttt{contains attempt at} directive and the
\texttt{intent\_rescue\_reason} contract.}
Many safety-critical criteria are phrased in terms of
\emph{what the agent attempted} rather than what succeeded.
A simulator bug in V8 prevented the \texttt{blood\_admin}
task type from executing, so well-formed orders for blood
products returned
$\{\texttt{status:error},\ \texttt{code:unknown\_task\_type}\}$
in the audit log. The V8 LLM judge correctly scored these as
agent intent (PASS); a naive world-state check would score
them as FAIL. To preserve this semantics deterministically
we added a \texttt{contains attempt at} directive to the
evaluator. An early implementation accepted \emph{any} audit
entry matching the tool/params filter regardless of
\texttt{result\_summary}, which turned out to be
over-permissive (\S F.4 below). The production directive
accepts an audit entry as evidence of attempt only if
\texttt{result\_summary == "ok"}, or
\texttt{result\_summary == "error"} with an
\texttt{error\_code} in a fixed simulator-side allowlist:
$\{\texttt{unknown\_task\_type},$
$\texttt{not\_implemented},$
$\texttt{simulator\_error},$
$\texttt{internal\_error},$
$\texttt{service\_unavailable}\}$.
Agent-side schema errors (\texttt{missing\_param},
\texttt{invalid\_params}, \texttt{unknown\_tool}) are not
valid attempts. In addition, every overlay entry using
\texttt{contains attempt at} must declare a free-text
\texttt{intent\_rescue\_reason} in the overlay YAML; the
orchestrator refuses to load the overlay if any entry is
missing this attestation. The attestation is the design
pattern; the allowlist is its enforcement.

\paragraph{Forensic walkthrough: SCJ-005-C04.}
A pre-tightening run of the overlay classified two trials
of SCJ-005-C04 (GPT-5.4 seeds $42$ and $44$) as
$\text{v9}=\text{PASS}$ where V8 had marked them FAIL, with
the disagreement classifier assigning the label
\texttt{infrastructure\_error}. On direct audit, however,
only one trial ($\_44\_t3$) carried a genuine sim-side
\texttt{unknown\_task\_type}. The other ($\_42\_t1$) contained
$12/13$ tool errors of type \texttt{missing\_param}: the
agent had omitted the required \texttt{details} field from
its \texttt{createClinicalOrder} calls and never once
produced a well-formed request. Under a principled reading
of ``attempted'' this trial had not attempted the order, it
had emitted malformed calls. Under the tightened directive,
$\_42\_t1$ correctly flips back to $\text{v9}=\text{FAIL}$
while $\_44\_t3$ retains its rescue. The pre-tightening
overlay was laundering an agent-side schema failure as a
clinical attempt. We report the walkthrough because the
failure mode---a naive intent check that accepts any error
code as evidence of intent---is easy to recreate, and the
fix (the sim-side allowlist plus per-entry attestation) is
narrow enough to be reused.

\paragraph{Agreement.}
Agreement was computed on
$n=264$ % CN:v9_n_observations
(criterion $\times$ trial) observations: $44$ overlay
criteria $\times$ $2$ models (Claude Opus 4.6, GPT-5.4)
$\times$ $3$ trials. The V8 judge verdict is the reference;
the V9 overlay verdict is the test. Overall raw agreement is
$76.1\%$ % CN:v9_agreement
with Cohen's
$\kappa = 0.402$ % CN:v9_kappa_overall
and
$6$ % CN:v9_safety_inversions
safety-critical verdict inversions. Per-category agreement
is given in Table~\ref{tab:v9_cat_agreement}.

\begin{table}[h]
\centering
\caption{V9 overlay vs.\ V8 judge agreement by category.
$n$ is (criteria~$\times$~trials). $\kappa$ is Cohen's;
PABAK is prevalence-and-bias-adjusted kappa. V8 PASS
prevalence is high across all categories, which is why
$\kappa$ is systematically lower than PABAK (the
$\kappa\ge 0.80$ discussion below).}
\label{tab:v9_cat_agreement}
\small
\begin{tabular}{lrrrrr}
\toprule
Category & $n$ & Agreement & V8 prev. & $\kappa$ & PABAK \\
\midrule
clinical\_communication     & $12$  & $58.3\%$ & $58.3\%$ & $0.211$ & $0.167$ \\
clinical\_reasoning         & $12$  & $75.0\%$ & $66.7\%$ & $0.526$ & $0.500$ \\
multi\_step\_workflows      & $78$  & $82.1\%$ & $78.2\%$ & $0.533$ & $0.641$ \\
safety\_critical\_judgment  & $54$  & $77.8\%$ & $83.3\%$ & $0.265$ & $0.556$ \\
temporal\_reasoning         & $108$ & $73.1\%$ & $82.4\%$ & $0.335$ & $0.463$ \\
\midrule
all                         & $264$ & $76.1\%$ & $79.5\%$ & $0.402$ & $0.523$ \\
\bottomrule
\end{tabular}
\end{table}

\paragraph{Disagreement taxonomy.}
A rule-based classifier
(\texttt{scripts/kappa\_validation.py::\_classify\_disagreement})
assigns one of six labels to each disagreeing observation in
priority order: \texttt{overlay\_wrong\_entity} (overlay
targets wrong tool/params); \texttt{infrastructure\_error}
(v9 PASS rescued by a sim-side error);
\texttt{judge\_hallucination} (v8 PASS with zero successful
matching tool calls in trajectory);
\texttt{intent\_execution\_split} (agent attempted but did
not complete; judge scored intent);
\texttt{vocab\_gap} (medication-class terminology mismatch);
and \texttt{conditional\_logic} (assertion has a conditional
clause the overlay cannot encode). Counts for $63$
disagreeing observations are in Table~\ref{tab:v9_labels}.
The dominant class is \texttt{judge\_hallucination} at
$73\%$ % CN:v9_judge_halluc_pct
($46$ of $63$): % CN:v9_judge_halluc_count
V8 marked the criterion PASS yet the trajectory contains
zero successful matching tool calls in the audit log. These
are judge false positives, and because most HealthCraft
safety-critical criteria concern whether a tool was invoked,
they are a direct ceiling on judge reliability.

\begin{table}[h]
\centering
\caption{Disagreement classifier counts over the $63$
disagreeing observations. Safety column counts the
safety-critical subset.}
\label{tab:v9_labels}
\small
\begin{tabular}{lrrr}
\toprule
Label & Count & \% of disagreements & Safety-critical \\
\midrule
judge\_hallucination       & $46$ & $73\%$ & $5$ \\
infrastructure\_error      & $5$  % CN:v9_infra_err_count
                                   & $8\%$  & $1$ \\
vocab\_gap                 & $5$  & $8\%$  & $0$ \\
intent\_execution\_split   & $4$  & $6\%$  & $0$ \\
overlay\_wrong\_entity     & $2$  & $3\%$  & $0$ \\
conditional\_logic         & $1$  & $2\%$  & $0$ \\
\midrule
total                      & $63$ & $100\%$ & $6$ \\
\bottomrule
\end{tabular}
\end{table}

\paragraph{Tiering.}
On the basis of the disagreement taxonomy each of the 44
overlay entries is assigned one of three tiers:

\begin{itemize}
  \item \textbf{Tier 1 (reward-safe): $10$ entries.} % CN:v9_tier1_count
  All six trials agree with the V8 judge. The overlay can
  replace \texttt{llm\_judge} for these criteria in reward
  computation without changing any verdict.
  \item \textbf{Tier 2 (research-only): $26$ entries.} % CN:v9_tier2_count
  Disagreement labels are a subset of $\{$\texttt{judge\_hallucination},
  \texttt{infrastructure\_error}$\}$. The overlay is
  plausibly correct and the judge is plausibly at fault,
  but we do not use the overlay for reward until physician
  adjudication has confirmed the direction.
  \item \textbf{Tier 3 (keep \texttt{llm\_judge}):
  $8$ entries.} % CN:v9_tier3_count
  At least one disagreement carries a label outside the
  judge-noise set (\texttt{overlay\_wrong\_entity},
  \texttt{intent\_execution\_split}, \texttt{vocab\_gap},
  \texttt{conditional\_logic}). These flag real overlay or
  semantic issues; the V8 judge remains binding until the
  overlay is corrected.
\end{itemize}

The tiering is presented not as a HealthCraft-specific
artefact but as a reusable template: given any
deterministic-overlay audit of an LLM judge, the six-label
classifier over disagreements induces a three-way tiering
(reward-safe~/ research-only~/ keep-judge), and only the
reward-safe tier changes the production reward signal.

\paragraph{Why $\kappa \ge 0.80$ is the wrong gate here.}
Cohen's $\kappa$ treats the reference grader as ground
truth. The disagreement taxonomy shows that treatment is
unsafe for this audit: $73\%$ of observed disagreements
trace to judge false positives, not overlay faults. A
perfect overlay audited against a judge with
$\sim 23\%$ false-positive rate on the audited subset
cannot reach $\kappa \ge 0.80$. This is an instance of the
kappa paradox in the high-prevalence regime (V8 PASS
prevalence
$79.5\%$ % CN:v9_v8_prev_overall
on the audited subset): agreement is dominated by the
shared positive class, and the chance-agreement
denominator grows faster than the observed-agreement
numerator. The honest statistic for
\emph{reliability-of-overlay-given-noisy-judge} is raw
agreement paired with the disagreement taxonomy; the
honest statistic for
\emph{reliability-of-judge-given-physician-adjudicated-truth}
is $\kappa$ against an adjudicated subset, which is
underway in a follow-up.

\paragraph{Prevalence correction: negative-class task slate.}
\label{app:v9-neg}
The kappa paradox in Table~\ref{tab:v9_cat_agreement}
is driven by a structural imbalance, not a measurement
artefact: the audited subset has a V8-PASS prevalence of
$79.5\%$, % CN:v9_v8_prev_overall
so even a well-calibrated overlay cannot reach
$\kappa\ge 0.80$ without first re-balancing what the
benchmark measures. The V8 task slate is dominated by
scenarios where the correct behaviour is to \emph{act}
(order the CT, start the antibiotic, escalate the
disposition); it under-samples scenarios where the
correct behaviour is to \emph{refrain} (no head CT for
a low-risk headache, no admission for a vasovagal
syncope with clean workup, no opioid for acute
mechanical back pain). The structural fix, shipped in
this release, is a $10$-task % CN:neg_tasks
negative-class slate committed to the task corpus in
which every criterion tests for restraint and is graded
by a \texttt{world\_state} predicate of the form
\texttt{audit\_log does NOT contain \ldots}. The slate
adds $82$ criteria % CN:neg_criteria
($14$ safety-critical), % CN:neg_safety_critical
distributed across four categories: four
safety-critical-judgment tasks (NEG-001 low-risk chest
pain, NEG-002 PECARN minor head trauma, NEG-003 simple
febrile seizure, NEG-004 asymptomatic severe
hypertension), three temporal-reasoning tasks (NEG-005
hypoglycaemic aphasia mimicking stroke, NEG-006 clean
serial chest-pain workup at hour four, NEG-007
alcoholic ketoacidosis distinguished from DKA), two
multi-step-workflow tasks (NEG-008 uncomplicated
pyelonephritis, NEG-009 acute low back pain without
red flags), and one clinical-reasoning task (NEG-010
benign vasovagal syncope). The post-NEG audit set is
$205$ tasks, % CN:post_neg_tasks
$2{,}337$ criteria, % CN:post_neg_criteria
and $529$ safety-critical. % CN:post_neg_safety_critical

\paragraph{Smoke-pilot results: task-distribution rebalance succeeds, criterion-prevalence correction does not.}
A 60-run smoke pilot (Claude Opus 4.7 and GPT-5.4, each at
$10$~tasks~$\times$~$3$~trials, seed=42,
\texttt{docs/NEG\_SMOKE\_PILOT.md}) establishes two facts.
First, the NEG slate achieves the task-level rebalance it
was designed for: mean Pass@1 is $0.500$ % CN:neg_smoke_claude_pass1
for Claude and $0.533$ % CN:neg_smoke_gpt_pass1
for GPT-5.4, compared to the V8 corpus means of $0.248$
(Claude) and $0.126$ (GPT-5.4). The benchmark now spans
$[0.00,\,1.00]$ Pass@1 with real representation at the
moderate-difficulty end rather than clustering at
$[0.15,\,0.30]$. Second, the criterion-level prevalence
hypothesis---that NEG criteria would average ${\sim}0.65$
criterion-PASS and pull the audit-subset prevalence into
the $0.55$--$0.75$ band---is \emph{falsified} by the same
pilot. Observed NEG-criterion PASS prevalence is
$0.929$ % CN:neg_smoke_crit_prev
(457 of 492 criterion-trials): restraint criteria
(``did~NOT~\ldots'') pass at
$0.962$, % CN:neg_smoke_restraint_prev
positive criteria (action required) pass at
$0.899$. % CN:neg_smoke_positive_prev
The structural reason is simple: a ``did~NOT~order~head~CT''
criterion is satisfied whenever the agent never orders a
head~CT, which is the modal frontier-model behaviour on a
PECARN low-risk scenario. Restraint-class criteria are
high-pass by construction whenever the agent has been
trained against the specific over-action the criterion
forbids. Pulling audit-subset prevalence into the target
band would instead require criteria in the $0.40$--$0.60$
PASS-rate regime, which neither pure-restraint nor
pure-action tasks reliably produce. The Appendix F target
band is therefore a task-level Pass@1 target, not a
criterion-level prevalence target; the kappa-paradox
mitigation that remains open is physician adjudication of
the Tier~2 subset, not further task authoring.

\paragraph{Safety yield: two exposed iatrogenic patterns.}
The pilot caught two clinically significant restraint
failures that the V8 corpus does not surface. On NEG-007
(alcoholic ketoacidosis with normal glucose), $4$ of $6$ % CN:neg_smoke_aka_insulin
trials across both models started an insulin infusion---a
pattern that produces iatrogenic hypoglycaemia because the
underlying acidosis is driven by starvation, not
hyperglycaemia. On NEG-005 (hypoglycaemic aphasia as a
stroke mimic), $2$ of $6$ % CN:neg_smoke_tpa_hypoglycemic
trials ordered tPA/alteplase before resolving the
hypoglycaemia---tPA in a hypoglycaemic patient is a
well-known wrong-answer pattern. These two failures are
the slate's primary safety yield. Three tasks (NEG-001,
NEG-003, NEG-006) were too easy, with Pass@1~$=1.00$ on
both models; they are candidates for revision or retirement
in a follow-up.

\begin{table}[h]
\centering
\caption{Post-NEG corpus: ex-ante projection vs observed
smoke-pilot result. $n$ is (criteria~$\times$~models~$\times$~trials).
The ``projected'' column is the ex-ante estimate this
subsection set out to validate; the ``observed'' column is
the smoke-pilot result and \emph{falsifies} it in every
category. Restraint criteria are trivially satisfied by
default behaviour in frontier models, so negative-class
criterion-PASS prevalence runs near the ceiling rather than
in the target band. The table is retained to document the
magnitude and direction of the miss.}
\label{tab:neg_prev}
\small
\begin{tabular}{lrrrrr}
\toprule
Category & $n_{\text{pre}}$ & prev$_{\text{pre}}$
         & $n_{\text{NEG}}$ & prev$_{\text{NEG}}$ (proj.\,/\,obs.)
         & prev$_{\text{post}}$ (proj.\,/\,obs.) \\
\midrule
safety\_critical\_judgment  & $54$  & $0.833$ & $192$ & $0.65\,/\,0.93$ & $0.69\,/\,0.91$ \\
temporal\_reasoning         & $108$ & $0.824$ & $144$ & $0.65\,/\,0.93$ & $0.72\,/\,0.88$ \\
multi\_step\_workflows      & $78$  & $0.782$ & $108$ & $0.65\,/\,0.93$ & $0.71\,/\,0.87$ \\
clinical\_reasoning         & $12$  & $0.667$ & $48$  & $0.65\,/\,0.93$ & $0.65\,/\,0.88$ \\
\midrule
all                         & $252$ & $0.805$ & $492$ & $0.65\,/\,0.93$ & $0.70\,/\,0.89$ \\
\bottomrule
\end{tabular}
\end{table}

\paragraph{Reproducibility.}
The full audit report (per-criterion verdicts, per-trial
labels, and the classifier inputs for every disagreement) is
committed at \texttt{docs/V9\_OVERLAY\_AUDIT.json}. The
validator script (\texttt{scripts/kappa\_validation.py})
takes no API dependencies and re-derives every number in
this appendix from the V8 trajectories at
\texttt{results/pilot-v8-\{claude-opus,gpt54\}/trajectories/}
and the overlay YAML. Running the validator against a clean
checkout reproduces \texttt{docs/V9\_OVERLAY\_AUDIT.json}.
The negative-class task files are at
\texttt{configs/tasks/*/task\_neg\_\{001..010\}\_*.yaml} and
pass the repository's preflight check
(\texttt{make preflight}) against the full schema and
runtime evaluator.
